\documentclass[12pt]{iopart}

\bibliographystyle{iopart-num}
\usepackage{cite}

%%\usepackage[numbers,square,sort&compress]{natbib}


\newcommand{\BibTeX}{Bib\TeX}
\newcommand{\REVTeX}{REV\TeX}

\sloppy 

\begin{document}
\nocite{*}

\title{The \texttt{iopart-num} \BibTeX{} style}

\noindent \qquad \\[-6pt] \qquad Version 2.2\\\qquad February 2, 2016

\author{M~A~Caprio}

\address{Department of Physics, University of Notre Dame,
Notre Dame, Indiana 46556-5670, USA}

\section{Introduction}

The \texttt{iopart-num} \BibTeX{} style is intended for use in
preparing manuscripts for Institute of Physics Publishing (IOPP) journals,
including Journal of Physics.  It provides numeric citation with
Harvard-like formatting, based upon the specifications in the IOPP author guidelines ``How to
prepare and submit an article for publication in an IOP Publishing journal using
\LaTeXe'' (February 2014) at \texttt{http://authors.iop.org}.
The \texttt{iopart-num} style is compatible with the \texttt{iopart}
document class, but it may be used with other document classes as well.

The \texttt{iopart-num} package is available on the Comprehensive
\TeX{} Archive Network (CTAN) as \texttt{/biblio/bibtex/contrib/iopart-num}.

\section{General instructions}

To use the \texttt{iopart-num} style, include the command
\verb+\bibliographystyle{iopart-num}+ in the document preamble.
Alternatively, if you wish to include the titles of journal articles
in the bibliographic entries, use
\verb+\bibliographystyle{iopart-num-long}+ (titles are optional in
most IOPP journals but required for some).

As usual for \BibTeX{}, the reference section is then inserted into
the document with the command \verb+\bibliography{...}+, where the
names of the necessary \BibTeX{} database files should be listed
between the braces.  Further general instructions on using \BibTeX{}
may be found in the \BibTeX{} documentation.

Some usage details depend upon whether or not \texttt{iopart-num} is
being used with the \texttt{natbib} package.  If your document is
being prepared with the \texttt{iopart} document class but
\textit{without} \texttt{natbib}, the section header for the
references must be manually inserted, with the command
\verb+\section*{References}+.  In this case, you may wish to use the
\texttt{cite} package to produce sorted and compressed lists of
references in the text.  If your document is being prepared with the
\texttt{iopart} document class and \textit{with} \texttt{natbib}, the
``References'' section header will be generated automatically.  You
may wish to load the \texttt{natbib} package with the options
\verb+numbers+, \verb+square+, and \verb+sort&compress+ to ensure
proper formatting of the references in the text, as
\verb+\usepackage[numbers,square,sort&compress]{natbib}+.

\section{Special bibliographic data fields}

\subsection{Journal sections}

Under IOPP style conventions, journal names should be set in italic
type.  However, for journals with multiple lettered sections, the IOPP
convention is that the journal section letter should appear in roman
type after the main journal name, \textit{e.g.}, ``\textit{J.\
Phys.\/} A''.  Most existing \BibTeX{} styles do not make special
provision for lettered sections.  Therefore, typically, the section
letter is either included as part of the journal name
\begin{verbatim}
  journal = "J. Phys. A",
  volume = "38",
\end{verbatim}
or as part of the volume number
\begin{verbatim}
  journal = "J. Phys.",
  volume = "A38",
\end{verbatim}
in the \BibTeX{} database entry.  The \texttt{iopart-num} style
instead introduces a new optional field \verb+section+ which can be
used to specify a journal section letter.  This section letter is set in
roman type.  Moreover, if the section letter already appears in
\textit{any} of the usual locations in the database entry (at the end
of the journal name, before the volume number, or after the volume
number),
\texttt{iopart-num} will recognize it and suppress its printing.
Therefore, when you are creating the
\BibTeX{} database entry for an article in a lettered journal section,
you can still include the section letter in the \verb+journal+ or
\verb+volume+ fields, for use with other \BibTeX{} styles, 
without adversely affecting the formatting for IOPP journals.  For
example, the entry for ref.~\cite{caprio2005:coherent} can be
generated with
\begin{verbatim}
  journal = "J. Phys. A",
  section = "A",
  volume = "38",
\end{verbatim}
or
\begin{verbatim}
  journal = "J. Phys.",
  section = "A",
  volume = "A38",
\end{verbatim}
or simply
\begin{verbatim}
  journal = "J. Phys.",
  section = "A",
  volume = "38",
\end{verbatim}
in the \BibTeX{} database entry.  Note that section names longer than a
single letter are also supported (\textit{e.g.},
``\textit{Phys. Rev.\/} ST Accel. Beams'').

\subsection{Journal issue numbers}

Journal issue numbers are not customarily included in references to
journal articles under the IOPP formatting conventions.
Therefore, the
\texttt{iopart-num} style ignores the
\verb+number+ field for articles.  
However, in some periodicals (such as popular magazines or certain
journal online supplements), pagination restarts from 1 with each
issue.  For such periodicals, the issue number is an essential part of
the bibliographic information needed to identify an article.  The
\texttt{iopart-num} style therefore supports an additional field
\verb+issue+ in the \BibTeX{} database entry, which can be used to 
enforce printing of the issue number.  If a value is specified for
\verb+issue+, this value is included included parenthetically after the volume
number in the reference, as in
ref.~\cite{zamfir2005:132te-beta-enam04}.

\subsection{Multivolume books}

The IOPP guidelines distinguish between a volume
in a {\it series}
and a volume of a {\it  multivolume book} or {\it set}.  For a volume in
a series, the series title and volume number are given in parentheses
after the book title~\cite{ex8}.
For an individual volume in a multivolume book, the book title is
given first, followed by the volume number and volume title~\cite{ex9}.

The
\texttt{iopart-num} style supports an additional field
\verb+volumetitle+ in the \BibTeX{} database entry, which can be used to 
specify the title for an individual volume of a multivolume
book, as in refs.~\cite{ex9,bohr1998:v2}.  For example, the entry for
ref.~\cite{bohr1998:v2} is generated with
\begin{verbatim}
  title = "Nuclear Structure",
  volume = 2,
  volumetitle = "Nuclear Deformations",
\end{verbatim}
in the \BibTeX{} database entry.  In contrast, most existing \BibTeX{} styles 
allow you to reference a volume of a multivolume
book by specifying \verb+title+ and
\verb+volume+, as in
refs.~\cite{ex7,siegbahn1965:v1}, but do not provide for inclusion of
any specific title for the individual volume.  A volume in a
series~\cite{ex8,iachello2006:liealg} is indicated in
\texttt{iopart-num}, as in most other \BibTeX{}
styles, 
by specifying
\verb+title+, \verb+volume+, and
\verb+series+ in the \BibTeX{} database entry.  

\subsection{E-prints, collaborations, and other \REVTeX{} data fields}

The \texttt{iopart-num} style supports several additional data fields
(\verb+archive+, \verb+collaboration+, \verb+eid+, \verb+eprint+, 
\verb+numpages+, and \verb+url+) introduced in
\REVTeX{}~4.


\section{Examples}

The entries in the reference list below provide examples of the
formatting of various types of references, of varying complexity,
including journal articles, articles in proceedings or collections,
books (individual, multivolume, or in a series), theses, and
unpublished references.  The
\BibTeX{} database entries used to generate these examples can be
found in the file \texttt{iopart-num.bib}.
Refs.~\cite{ex1,ex2,ex3,ex4,ex5,ex6,ex7,ex8,ex9} are based upon example entries
from the IOPP guidelines.

\section*{References}
\bibliography{iopart-num}

\end{document}

